---
title: "Practical Machine Learning - Course Project"
author: "Bharat Dogra"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This report outlines the process of predicting the manner in which participants performed barbell lifts using data from accelerometers. The goal was to predict the `classe` variable in the training set using a machine learning model trained on wearable sensor data.

## Load Required Libraries

```{r libraries}
library(caret)
library(randomForest)
```

## Load and Clean the Data

```{r load-data}
train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(train_url, na.strings = c("NA", "#DIV/0!", ""))
testing  <- read.csv(test_url, na.strings = c("NA", "#DIV/0!", ""))

# Remove columns with mostly NA values
training <- training[, colSums(is.na(training)) == 0]

# Remove non-predictor columns (IDs, timestamps)
training <- training[, -c(1:7)]

# Make sure the target variable is a factor
training$classe <- as.factor(training$classe)

# Clean the test set to match training columns
testing <- testing[, names(training)[-ncol(training)]]
```

## Train the Random Forest Model

```{r train-model}
set.seed(12345)
modelFit <- train(classe ~ ., data = training, method = "rf",
                  trControl = trainControl(method = "cv", number = 3),
                  ntree = 100)
```

## Evaluate Accuracy on Training Set

```{r evaluate-training}
train_predictions <- predict(modelFit, newdata = training)
confusionMatrix(train_predictions, training$classe)
```

## Predict on Final Test Set (20 Cases)

```{r predict-final}
final_predictions <- predict(modelFit, newdata = testing)
final_predictions
```

## Formatted Answers for Submission

```{r quiz-answers}
for (i in 1:length(final_predictions)) {
  cat(paste("Problem", i, ":", final_predictions[i]), "\n")
}
```

## Summary

The Random Forest model was chosen for its robustness and high accuracy in classification problems. After data cleaning and model training, the accuracy on the training set was approximately 100%. The model was then used to generate predictions for the final 20 test cases. These predictions are ready for submission to the Course Project Prediction Quiz.



